---
title: "Ada & Archibald MacLeish Field Station Data"
author: "Ben Baumer, Rose Goueth, Wencong Li (Priscilla) and Weijia Zhang (Vega), Paul Wetzel"
date: "04-10-2016"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using MacLeish Weather Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(knitr)
opts_chunk$set(fig.width = 6, fig.height = 4)
```

## MacLeish

The [`macleish`](http://www.github.com/beanumber/macleish) package for R contains weather data collection at the [Ada & Archibald MacLeish Field Station](http://www.smith.edu/ceeds/macleish.php) in Whately, MA, as well as spatial data detailing various features of the property. This vignette will illustrate how these data can be used. 

### Getting started:
To start using the package, load the `macleish` package in to your R session:

```{r, message=FALSE}
library(macleish)
```


## Using the MacLeish Field Station weather data

Two data frames are included in the package: `whately_2015` and `orchard_2015`. These contain weather data averaged over every ten minute period for the calendar year 2015. The `head` and `tail` will give you the first (last) 6 readings from the whately weather station from 2015, respectively.

```{r}
head(whately_2015)
tail(whately_2015)
```

```{r, eval=FALSE}
head(orchard_2015)
tail(orchard_2015)
```

To access data during wider time spans, make use of the [`etl` framework](http://www.github.com/beanumber/etl) to download the data and import it into a database. Please see the documentation for `etl_extract` for further details and examples. 

```{r, eval=FALSE}
help("etl_extract.etl_macleish")
```


### The 2015 MacLeish weather report

This portion of the vignette is largely a re-creation of the Macleish Field Station 2015 Weather Report. You can download [the orginal report](http://www.smith.edu/ceeds/documents/DataInformation2015.pdf) as a PDF file.]

#### Introduction to MacLeish Weather Stations
Weather conditions at Smith College's [Ada & Archibald MacLeish Field Station](http://www.smith.edu/ceeds/macleish.php) are recorded at the **WhatelyMet** weather station. This weather collection site is located at the end of Poplar Hill Road in Whately, Massachusetts, USA (Latitude 42 deg. 26 min. 56 sec. N Longitude 72 deg. 40 min. 50 sec. W). The meteorological instruments of WhatelyMet (except the rain gauge) are mounted at the top of a tower 25.3 m tall, well above the surrounding forest canopy. The tower is located on a local ridge at an elevation 250.75m above sea level.

> Get pictures from Paul!

The second weather data collection site, known as **OrchardMet**, is located about 250 m north of the first tower in an open field next to an apple orchard. Full canopy trees (~20 m tall) are within 30 m of this station. This station has a standard instrument configuration with temperature, relative humidity, solar radiation, and barometric pressure measured between 1.5 and 2.0 m above the ground. Wind speed and direction are measured on a 10 m tall tower and precipitation is measured on the ground. Ground temperature is measured at 15 and 30 cm below the ground surface 2 m south of the tower. The tower is located 258.1 m above sea level. Data collection at OrchardMet began on June 27th, 2014. 

This documentation describes the weather data collected at WhatelyMet and OrchardMet during 2015. It provides highlights of the year's weather as well as general trends that were observed. We hope that you find the doucumentation both interesting and useful. All data reported in this document are available through the [`macleish`](http://www.github.com/beanumber/macleish) package for R. The following will give you an outline of how to use the package by showing demos of the 2015 weather data. 

#### Variables used in this doucumentation:

* Temp_C_Avg: the average temperature (degrees C)
* WSpd_mps: Average windspeed (meters per second)
* Wdir_deg: Wind direction (degree)
* RH_per_Avg: Relative Humidity (%)
* Press_mb_Avg: Average atomspheric pressure (mb)
* SlrW_Avg: WhatelyMet Average Solar Radiance (W)
* Rain_mm_Tot: Main Rainfall amount per day (mm)
* PAR_Den_Avg: OrchardMet Measures solar radiation (W)
* PAR_Tot_Avg: OrchardMet Measures total solar radiation (W)

#### Definitions of Climatic Indicators 
The precipitation and biological temperature indicators calculated for this doucumentation provide a method of comparing general, biologically meaningful climatic trends in different years. The indicators are based on average daily temperatures and rainfall summations. Even though days are in the name of the temperature indicators, they do not actually reflect periods of 24 hours. The indicators really measure the number of degrees the average temperature was above or below a threshold temperature each day of the year. These values are described at [Nelson Institute Center for Climate Research Webpage](http://ccr.aos.wisc.edu/climate_modeling/wisconsin_Climate/descriptions.php). All temperature terms were measured in degrees Celsius. 

> What? I don't understand what this is saying.

### Example 1: Air Temperature & Relative Humidity
In this example we visualize the trend of average daily temperature at the MacLeish Field Station over 2015.

```{r daily, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
daily <- whately_2015 %>%
  mutate(the_date = as.Date(when)) %>%
  group_by(the_date) %>%
  summarize(N = n(), avg_temp = mean(Temp_C_Avg), 
            max_temp = max(Temp_C_Avg), min_temp = min(Temp_C_Avg))
```


```{r temp-plot, message=FALSE}
ggplot(data = whately_2015, aes(x = when, y = Temp_C_Avg)) +
  geom_line(color = "lightgray") + 
  geom_smooth() + 
  geom_line(data = daily, aes(x = as.POSIXct(the_date), 
                              y = avg_temp)) + 
  xlab(NULL) + ylab("Temperature (Celsius)")
```


As shown in the table below, the monthly average temperature in 2015 ranged from -6.4 °C in January to 21.6°C in July. 

```{r temp-table}
monthly_w <- whately_2015 %>%
  mutate(the_month = month(when, label = TRUE, abbr = FALSE)) %>%
  group_by(the_month) %>%
  summarize(avg_temp = mean(Temp_C_Avg))
monthly_o <- orchard_2015 %>%
  mutate(the_month = month(when, label = TRUE, abbr = FALSE)) %>%
  group_by(the_month) %>%
  summarize(avg_temp = mean(Temp_C_Avg))
monthly_w %>%
  inner_join(monthly_o, by = "the_month")
```

The highest recorded temperature at WhatelyMet was 31.6 °C on July 23th. The lowest recorded temperature was -21.4 °C on January 6th. 

> These don't seem to be correct. What is going on? 

```{r temp-extremes}
daily %>%
  arrange(desc(max_temp)) %>%
  head()
daily %>%
  arrange(min_temp) %>%
  head()
```

The average daily temperature range recorded from both stations (daily maximum-minimum) was -10.5 °C. The maximum daily temperature range was 20.8 °C at both stations but occurred on May 4th at OrchardMet and on May 8th at WhatleyMet. The daily minimum temperature range was 2.3 °C and occurred on January 26th. However daily temperature ranges of less than three degrees also occurred on February 9th, April 9th, June 1st, and June 28th. 

> Again, these numbers don't seem to match up. 

```{r temp-range}
daily %>%
  mutate(temp_range = max_temp - min_temp) %>%
  summarize(avg_temp_range = mean(temp_range), 
            max_temp_range = max(temp_range))
daily %>%
  mutate(temp_range = max_temp - min_temp) %>%
  arrange(temp_range) %>%
  head()
```

Percent average monthly relative humidity ranged from 56.4% (March) to 94% (December). Interestingly, WhatelyMet relative humidity values were notably higher than OrchardMet values starting in August. It is unclear why this pattern occurred during the last part of the year.


```{r humidity-table}
monthly_w <- whately_2015 %>%
  mutate(the_month = month(when, label = TRUE, abbr = FALSE)) %>%
  group_by(the_month) %>%
  summarize(avg_humidity = mean(RH_per_Avg))
monthly_o <- orchard_2015 %>%
  mutate(the_month = month(when, label = TRUE, abbr = FALSE)) %>%
  group_by(the_month) %>%
  summarize(avg_humidity = mean(RH_per_Avg))
monthly_w %>%
  inner_join(monthly_o, by = "the_month")
```

### Example 3: Wind

Wind speed and direction is measured every second and then averaged and logged every 10 minutes at both weather stations. OrchardMet measures wind midway up the forest canopy (10 m high) in a small woodland clearing while WhatleyMet measures wind 25.3 m above the ground which is above the tree canopy.

```{r, message=FALSE, warning=FALSE}
require(clifro)
orchard_2015 %>%
  with(windrose(WSpd_mps, Wdir_deg))
```

OrchardMet is located in a woodland clearing where winds are buffeted by trees of the forest canopy. Most wind comes from the west and southwest, with 67% of the time less than 1 m/s. Wind speeds did not average over 3 m/s at any time of the year. 

```{r, message=FALSE}
whately_2015 %>%
  with(windrose(WSpd_mps, Wdir_deg))
```

WhatleyMet is above the forest canopy and recorded winds frequently from the northwest, west, and south during the year. Average wind velocities were higher, with 46% of the measurements falling between 1 and 2 m/s, with maximum average velocities up to 6 m/s. 


### Example 5: Precipitation

The `when` variable records a timestamp for each row of the `macleish` weather data. The observations record the average of each measurement over a 10 minute interval. 

A total of 1306 mm of precipitation was recorded at WhatelyMet during 2015. 

```{r}
sum(whately_2015$Rain_mm_Tot)
```

Although precipitation is relatively evenly distributed across the year, precipitation totals were highest in June and lowest in March.

```{r rain-table}
monthly_w <- whately_2015 %>%
  mutate(the_month = month(when, label = TRUE, abbr = FALSE)) %>%
  group_by(the_month) %>%
  summarize(total_precip = sum(Rain_mm_Tot))
monthly_o <- orchard_2015 %>%
  mutate(the_month = month(when, label = TRUE, abbr = FALSE)) %>%
  group_by(the_month) %>%
  summarize(total_precip = sum(Rain_mm_Tot))
monthly_w %>%
  inner_join(monthly_o, by = "the_month")
```

February and March had the greatest number of precipitation events. Annual precipitation accumulated at a low, but fairly steady rate through June. After June rainfall levels increased and total accumulation increased sharply to the end of the calendar year. Starting in June a series of large rainfall events (40-50 mm) occurred regularly over the summer. The largest precipitation event in 2015 occurred on September 30th when 113.8 mm (8.7% of the annual total) of precipitation fell. During 2015, there were 15, 25.4 mm (1”) precipitation events; 2, 50.8 mm (2”) precipitation events; and 1, 76.2 mm (3”) precipitation event. 



Since the data was recorded every ten minutes, we need to group the data daily in order to get the mean rainfall amount per day.

```{r, message=FALSE}
daily_precip <- whately_2015 %>%
  mutate(the_date = as.Date(when)) %>%
  group_by(the_date) %>%
  summarize(N = n(), total_precip = sum(Rain_mm_Tot)) %>%
  mutate(cum_precip = cumsum(total_precip),
         cum_rescale = (cum_precip / max(cum_precip)) * max(total_precip))
```

The following plot will show you how much rainfall you would have per day at macleish field station over 2015.

```{r daily-precip, message=FALSE}
ggplot(data = daily_precip, 
       aes(x = the_date, y = total_precip)) +
  geom_bar(stat = "identity") + 
  geom_line(aes(y = cum_rescale), color = "blue") + 
  ylab("Daily Precipitation (mm)") + xlab(NULL)
```

## Using the Macleish Field Station Geographic Data

### Example 6: Maps

We can create a map of MacLeish using the `leaflet` package and the spatial data in the `macleish_layers` object. This is a `list` of `Spatial*DataFrame` objects that contain various pieces of information about the field station, including the location of buildings, streams, trails, and landmarks.

```{r}
names(macleish_layers)
```

These layers can be combined to provide a fuller picture of the activity at the Field Station. 

```{r, message=FALSE}
library(leaflet)
map <- leaflet() %>%
  addTiles() %>%
  addPolygons(data = macleish_layers[["boundary"]],
              weight = 1) %>%
  addMarkers(data = macleish_layers[["landmarks"]], 
             popup = ~Label)
map
```

